---
title: "Labroratory work №1. Research of data by regions of Russian Federation. Variant №8"
author: "Andrey Vladimirov"
date: "`r format(Sys.Date(), '%b %d %y')`"
output: word_document
---

```{r setup, include = FALSE}
library(knitr)
library(nortest)  #library for Anderson-Darling test
library(corrplot) #library for visualization of correlation matrices
library(Hmisc)  #library to calculate the correlation matrix
knitr::opts_chunk$set(echo = FALSE)
```

# SECTION 0. Uploading data

```{r import}
# set the name of the document
file_name = 'Data_Vladimirov.csv'
# uploading the table with observations from csv file with data (in project directory, data for save to RData file) 
regions_data_load <- read.csv2(file_name, sep = ',', encoding = 'UTF-8', check.names = FALSE)
# data for analysis
regions_data <- regions_data_load
# make Federal district as factors 
regions_data$Federal_district <- factor(regions_data$Federal_district)
# delete first column with Regions names
regions_data <- regions_data[, c(-1)]
# throw out the missed values
regions_data <- na.omit(regions_data)
```

Full list of data you can find at the file `r file_name`, located in the project directory.

Below are the first five values from the table.

```{r print_list}
# print first five rows of list
kable(head(regions_data))
```

File contains `r nrow(regions_data)` observations and `r ncol(regions_data)` variables.

### Designations:

*   Regions - full list of the Russian Federation region;

*   Federal district - federal district of the region;

*   Y - retail trade turnover per capita in 2016;

*   X1 - the number of small businesses per 10000 people in 2016;

*   X2 - the use of information and communication technologies in organizations:           personal computers in 2016;

*   X3 - expenditures of the consolidated budgets of the constituent entities of           the Russian Federation: total in 2015;

*   X4 - real cash income of the population in 2016.

# SECTION 1. Working with variables

### Description:

In this section need to do the following actions:

  1) find actual distribution densities of variables
with superimposed theoretical normal law densities,
and place graphs on one canvas;

  2) perform the Shapiro-Wilk test for normality:
a table with test statistics, P-value and output 
for each variable;

  3) construct cross scatter point plots,
make markers semi-transparent;

  4) Study the relationship of indicators: graphical 
representation of the correlation matrix, hide 
insignificant correlation coefficients, 
show the rest with squares.

### Implementation:

#### First step

Calculate factors descriptive statistics, standard deviations and coefficients of variation.

General descriptive statistics (function 'summary'):
```{r descr_stat}
# print general descriptive statistics
kable(summary(regions_data[2:6]))
```

Table with mean, standard deviations and coefficients of variation:
```{r mean_data}
# vector with mean values
mean_data <- round(apply(regions_data[, 2:6], 2, mean), 2)
# vector with standard deviations values
stand_dev <- round(apply(regions_data[, 2:6], 2, sd), 2)
# vector with coefficients of variation values
coeff_var <- round(((stand_dev / mean_data) * 100), 2)
# create table with values of descriptive statistics
desc_table <- rbind(mean_data, stand_dev, coeff_var)
# names of rows desc_table
row.names(desc_table) <- c('Mean', 'Standard deviations', 'Coefficient of variation')
# print table wit values of descriptive statistics
kable(desc_table)
```

**Conclusion**: the coefficients of variation show that the data is heterogeneous, as some percentages are greater than 20 percent.

```{r distr_plots}
# split the canvas into six values
par(mfrow = c(2, 3))
# setting external fields
par(oma = c(0, 0, 1.5, 0))
# setting internal margins
par(mar = c(4, 4, 0.5, 0.5))
# for each indicator
for (iter in 2:6) {
  # write the column into a separate variable
  x <- regions_data[, iter]
  # draw a histogram
  hist(x, 
       freq = FALSE, col = 'wheat', 
       xlab = colnames(regions_data)[iter], 
       ylab = 'Плотность распр.', main = '')
  # actual distribution density line
  lines(density(x), col = 'red', lwd = 2)
  # theoretical density line
  curve(dnorm(x, mean = mean(x), sd = sd(x)), 
        col = 'darkblue', lwd = 2, add = TRUE)
}
# plots title
title(main = 'Гистограммы распределения показателей', 
      outer = TRUE, cex = 1.5)
# return settings
par(mfrow = c(1, 1))
```

**Conclusion**: distributions of all indicators are asymmetric compared to the normal law.

#### Second step

```{r shapiro_wilk_test}
# test statistics
test_stat <- sapply(regions_data[, 2:6], function(x) {
  round((shapiro.test(x)$statistic), 2)
})
# structure of object
p_value <- sapply(regions_data[, 2:6], function(x) {
  round((shapiro.test(x)$p.value), 2)
})
# print table with values
kable(data.frame(p_value, test_stat))
```

**Conclusion**: the hypothesis of normal distribution is rejected, since all factors have P-values less than the significance level (0.05 or 5%).

#### Third step
```{r cross_scatter_plots}
# cross-scatter plots
pairs(regions_data[, -1], pch = 1,
      col = rgb(1, 0, 0, alpha = 0.4),
      bg = rgb(0, 0, 1, alpha = 0.4), cex = 1.1)
```

**Conclusion**: strict straight linear relationship between Y and X1, Y and X3, X1 and X3, and strict reverse relationship between X2 and X4.

#### Fourth step

```{r corr_matrix}
# coefficient Pearson and P-value
corr_data <- rcorr(as.matrix(regions_data[, -1]))
# print plot with correlation matrix
corrplot(corr_data$r, order = 'original', diag = FALSE,
         insig = 'blank', p.mat = corr_data$p, 
         sig.level = 0.05, method = 'square')
# title of plot
title(ylab = 'Матрица корреляции')
```

**Conclusion**: meaningful connection between Y and X1, Y and X3, X1 and X3, X2 and X4.

# SECTION 2

### Description:

In this section we need to do the same actions as in the first section, but with the logarithm data.

### Implementation:

```{r make_data_log}
# copy data before logarithm
regions_data_log <- regions_data
# logarithm values and round
regions_data_log[, 2:6] <- round(apply(regions_data[, 2:6], 2, log), 2)
# print first five rows of list
kable(head(regions_data_log))
```

#### First step

Calculate factors descriptive statistics, standard deviations and coefficients of variation.

General descriptive statistics (function 'summary'):
```{r descr_stat_log}
# print general descriptive statistics
kable(summary(regions_data_log[2:6]))
```

Table with mean, standard deviations and coefficients of variation:
```{r mean_data_log}
# vector with mean values
mean_data_log <- round(apply(regions_data_log[, 2:6], 2, mean), 2)
# vector with standard deviations values
stand_dev_log <- round(apply(regions_data_log[, 2:6], 2, sd), 2)
# vector with coefficients of variation values
coeff_var_log <- round(((stand_dev_log / mean_data_log) * 100), 2)
# create table with values of descriptive statistics
desc_table_log <- rbind(mean_data_log, stand_dev_log, coeff_var_log)
# names of rows desc_table_log
row.names(desc_table_log) <- c('Mean', 'Standard deviations', 'Coefficient of variation')
# print table wit values of descriptive statistics
kable(desc_table_log)
```

**Conclusion**: the coefficients of variation show that the data is homogeneous, as some percentages are less than 20 percent.

```{r distr_plots_log}
# split the canvas into six values
par(mfrow = c(2, 3))
# setting external fields
par(oma = c(0, 0, 1.5, 0))
# setting internal margins
par(mar = c(4, 4, 0.5, 0.5))
# for each indicator
for (iter in 2:6) {
  # write the column into a separate variable
  x <- regions_data_log[, iter]
  # draw a histogram
  hist(x, 
       freq = FALSE, col = 'wheat', 
       xlab = colnames(regions_data_log)[iter], 
       ylab = 'Плотность распр.', main = '')
  # actual distribution density line
  lines(density(x), col = 'red', lwd = 2)
  # theoretical density line
  curve(dnorm(x, mean = mean(x), sd = sd(x)), 
        col = 'darkblue', lwd = 2, add = TRUE)
}
# plots title
title(main = 'Гистограммы распределения показателей', 
      outer = TRUE, cex = 1.5)
# return settings
par(mfrow = c(1, 1))
```

**Conclusion**: distributions of all indicators are asymmetric compared to the normal law.

#### Second step

```{r shapiro_wilk_test_log}
# test statistics
test_stat_log <- sapply(regions_data_log[, 2:6], function(x) {
  round((shapiro.test(x)$statistic), 2)
})
# structure of object
p_value_log <- sapply(regions_data_log[, 2:6], function(x) {
  round((shapiro.test(x)$p.value), 2)
})
# print table with values
kable(data.frame(p_value_log, test_stat_log))
```

**Conclusion**: the hypothesis of normal distribution is rejected, since Y, X1, X2, X4 factors have P-values less than the significance level (0.05 or 5%). X3 have P-value greater than significance level(0.05 or 5%), therefore we do not reject the hypothesis.

#### Third step
```{r cross_scatter_plots_log}
# cross-scatter plots
pairs(regions_data_log[, -1], pch = 1,
      col = rgb(1, 0, 0, alpha = 0.4),
      bg = rgb(0, 0, 1, alpha = 0.4), cex = 1.1)
```

**Conclusion**: strict straight linear relationship between Y and X1, Y and X3, X1 and X3, and strict reverse relationship between X2 and X4.

#### Fourth step

```{r corr_matrix_log}
# coefficient Pearson and P-value
corr_data_log <- rcorr(as.matrix(regions_data_log[, -1]))
# print plot with correlation matrix
corrplot(corr_data_log$r, order = 'original', diag = FALSE,
         insig = 'blank', p.mat = corr_data_log$p, 
         sig.level = 0.05, method = 'square')
# title of plot
title(ylab = 'Матрица корреляции')
```

**Conclusion**: meaningful connection between Y and X1, Y and X3, X1 and X3, X2 and X4.

```{r save_data}
# save data to RData file
save(list = c('regions_data_load', 'regions_data', 'regions_data_log'),
     file = 'labs_VladimirovAA_data.RData')
# save image to RData file
save.image(file = 'labs_VladimirovAA_data.RData')
```